Adaptive Approximation for Multivariate Linear Problems \\ with Inputs Lying in a Cone

### 2018-11-07 RICAM2018NovTalk.tex
revising

	\newcommand{\scoop}[1]{\parbox{#1}{\includegraphics[width=#1]{IceCreamScoop.eps}}\xspace}
	\newcommand{\smallscoop}{\scoop{1cm}}
	\newcommand{\medscoop}{\parbox{1.8cm}{\includegraphics[width=1.8cm]{IceCreamScoop.eps}}\xspace}
	\newcommand{\largescoop}{\scoop{3cm}}
	\newcommand{\ICcone}[1]{\parbox{#1}{\includegraphics[width=#1,angle=270]{MediumWaffleCone.eps}}\xspace}
	\newcommand{\medcone}{\ICcone{1.2cm}}
	\newcommand{\largercone}{\parbox{2.2cm}{\vspace*{-0.2cm}\includegraphics[width=1cm,angle=270]{MediumWaffleCone.eps}}\xspace}
	\newcommand{\largecone}{\ICcone{1.8cm}}
	\newcommand{\smallcone}{\parbox{0.65cm}{\includegraphics[width=0.3cm,angle=270]{MediumWaffleCone.eps}}\xspace}

@@

\begin{frame}{Context for This Talk}

\centerline{
    \begin{tabular}{rp{0.4\textwidth}@{\qquad}p{0.4\textwidth}}
    \alert{Henryk} & Many nice tractability results for \medscoop &
    May we obtain analogous results for \medcone \hspace{7ex} ? \tabularnewline
    \alert{Houman} & Let's learn the appropriate kernel from the function data & This will only work for nice functions in a \medcone
    \end{tabular}}
\end{frame}

@@

    \item<2>[Step 3] Let $\tf \in \cf$ be the \alert{minimum} norm interpolant of the data $L_1(f), \ldots, L_n(f)$

    \item[Step 4]  \uncover<2>{If $C\bignorm[\cf]{\tf} \le \rho$ for some preset \alert{inflation} factor, $C$, }then return $A(f,\varepsilon) = \Sapp(f,n)$\uncover<2>{; otherwise, choose $\rho = 2C\bignorm[\cf]{\tf}$, and go to Step 2}
\end{description}

@@

\uncover<2>{This succeeds for the \alert{cone} \smallcone defined as those functions in $\cf$ whose norms are not much larger than their minimum norm interpolants.   \hyperlink{ConeFrame}{\beamerreturnbutton{Back}}}

@@

\begin{align*}
    \only<1>{\cf &:= \left \{ f = \sum_{\vj \in \cn} \hf(\vj) u_{\vj} : \norm[\cf]{f} := \norm[2]{\left(\frac{\bigabs{\hf(\vj)}}{\lambda_{\vj}} \right)_{\vj \in \cn}} \right \} \qquad \begin{minipage}{3cm}\raggedright\alert{$\vlambda$ affects convergence rate \& tractability}\end{minipage}\\
    \cg &: = \biggl \{ g = \sum_{\vj \in \cn} \hg(\vj) v_{\vj} : \norm[\cg]{g} := \bignorm[2]{\hg}\biggr \}, \qquad v_{\vj} = S(u_{\vj}) \\[-2ex]}
      & \alert{\lambda_{\vj_1} \ge \lambda_{\vj_2} \ge \cdots}, \qquad
      n_0 < n_1 < n_2 < \cdots,  \qquad
 	\sigma_k(f) = \sum_{i=n_{k-1}+1}^{n_k}  \bigabs{\hf(\vj_i)}, \quad k \in \naturals \\
     \cc &: = \left\{f \in \cf :  
 	 \sigma_\ell(f) \le a b^{\ell - k}\sigma_k(f) \quad \forall k \le \ell \right\} \quad a > 1 > b \quad \text{\alert{series coef.\ decay steadily}}\\
 	 \Sapp(f,n_k) &= \sum_{i=1}^{n_k} \hf(\vj_i) v_{\vj_i} \text{ is optimal for fixed }n_k, \qquad \norm[\cg]{S(f) - \Sapp(f,n_k)} \le \frac{ab \sigma_k(f)}{\sqrt{1 - b^2}} \\[-1ex]
 	 \alert{A(f,\varepsilon)} & \alert{= \Sapp(f,n_k) \text{ for the smallest $k$ satisfying } \sigma_k(f) \le \frac{\varepsilon \sqrt{1 - b^2}}{ab}}
 	 \only<2>{\\ 
 	 \MoveEqLeft{\alert{\cost(A,\cc,\varepsilon,\rho) = n_{\ell^\dagger}, \quad 
 	 \ell^\dagger \le \min \left \{\ell \in \naturals : \frac{\rho^2}{\varepsilon^2} \le \frac{(1 - b^2)}{a^2b^2} \left[ \sum_{k=1}^{\ell-1} \frac{b^{2(k-\ell)}}{a^2\lambda_{n_{k-1}+1}^2} + \frac{1}{\lambda_{n_{\ell-1}+1}^2}\right]   \right\}}} \\
 	 \MoveEqLeft{\alert{\cost(A,\cc,\varepsilon,\rho) \text{ essentially no worse than } \comp(\ca(\cc,\LambdaAll),\varepsilon,\rho)}}}
\end{align*}

@@

    \begin{align*}
    \only<1>{\cf &:= \left \{ f = \sum_{\vj \in \natzero^d} \hf(\vj) u_{\vj} : \norm[\cf]{f} :=  \norm[\infty]{\left(\frac{\bigabs{\hf(\vj)}}{\lambda_{\vj}} \right)_{\vj \in \cn}} \right \} \qquad \begin{minipage}{3cm}\raggedright\alert{$\vlambda$ affects convergence rate \& tractability}\end{minipage}\\
    \cg &: = \biggl \{ g = \sum_{\vj \in \cn} \hg(\vj) v_{\vj} : \norm[\cg]{g} := \bignorm[1]{\hg}\biggr \}, \qquad v_{\vj} = S(u_{\vj}) = u_{\vj} \\[-2ex]}
    \lambda_{\vj} & = \Gamma_{\norm[0]{\vj}} \prod_{\substack{\ell = 1 \\ j_{\ell} > 0}}^d \gamma_\ell s_{j_\ell} \quad \text{\alert{product, order, and smoothness dependent weights}} \\
      & \alert{\lambda_{\vj_1} \ge \lambda_{\vj_2} \ge \cdots}, \qquad
      \Sapp(f,n) = \sum_{i=1}^{n} \hf(\vj_i) u_{\vj_i} ,  \qquad
 	\sigma_k(f) = \sum_{i=n_{k-1}+1}^{n_k}  \bigabs{\hf(\vj_i)}, \quad k \in \naturals \\
     \cc &: = \left\{f \in \cf :  
 	 \sigma_\ell(f) \le a b^{\ell - k}\sigma_k(f) \quad \forall k \le \ell \right\} \quad a > 1 > b \quad \text{\alert{series coef.\ decay steadily}}\\
 	 \Sapp(f,n_k) &= \sum_{i=1}^{n_k} \hf(\vj_i) v_{\vj_i} \text{ is optimal for fixed }n_k, \qquad \norm[\cg]{S(f) - \Sapp(f,n_k)} \le \frac{ab \sigma_k(f)}{\sqrt{1 - b^2}} \\[-1ex]
 	 \alert{A(f,\varepsilon)} & \alert{= \Sapp(f,n_k) \text{ for the smallest $k$ satisfying } \sigma_k(f) \le \frac{\varepsilon \sqrt{1 - b^2}}{ab}}
 	 \only<2>{\\ 
 	 \MoveEqLeft{\alert{\cost(A,\cc,\varepsilon,\rho) = n_{\ell^\dagger}, \quad 
 	 \ell^\dagger \le \min \left \{\ell \in \naturals : \frac{\rho^2}{\varepsilon^2} \le \frac{(1 - b^2)}{a^2b^2} \left[ \sum_{k=1}^{\ell-1} \frac{b^{2(k-\ell)}}{a^2\lambda_{n_{k-1}+1}^2} + \frac{1}{\lambda_{n_{\ell-1}+1}^2}\right]   \right\}}} \\
 	 \MoveEqLeft{\alert{\cost(A,\cc,\varepsilon,\rho) \text{ essentially no worse than } \comp(\ca(\cc,\LambdaAll),\varepsilon,\rho)}}}
\end{align*}

@@

@@

    \item<2>[Step 3] Let $\tf \in \cf$ be the \alert{minimum} norm interpolant of the data $L_1(f), \ldots, L_n(f)$

    \item[Step 4]  \uncover<2>{If $C\bignorm[\cf]{\tf} \le \rho$ for some preset \alert{inflation} factor, $C$, }then return $A(f,\varepsilon) = \Sapp(f,n)$\uncover<2>{; otherwise, choose $\rho = 2C\bignorm[\cf]{\tf}$, and go to Step 2}
\end{description}

\uncover<2>{This succeeds for the \alert{cone} \smallcone defined as those functions in $\cf$ whose norms are not much larger than their minimum norm interpolants.   \hyperlink{ConeFrame}{\beamerreturnbutton{Back}}}

>>>>

	\newcommand{\scoop}[1]{\parbox{#1}{\includegraphics[width=#1]{IceCreamScoop.eps}}\xspace}
	\newcommand{\smallscoop}{\scoop{1cm}}
	\newcommand{\medscoop}{\scoop{1.8cm}}
	\newcommand{\largescoop}{\scoop{3cm}}
	\newcommand{\ICcone}[1]{\parbox{#1}{\includegraphics[width=#1,angle=270]{MediumWaffleCone.eps}}\xspace}
	\newcommand{\medcone}{\ICcone{1.2cm}}
	\newcommand{\largercone}{\parbox{2.2cm}{\vspace*{-0.2cm}\includegraphics[width=1cm,angle=270]{MediumWaffleCone.eps}}\xspace}
	\newcommand{\largecone}{\ICcone{1.8cm}}
	\newcommand{\smallcone}{\parbox{0.65cm}{\includegraphics[width=0.3cm,angle=270]{MediumWaffleCone.eps}}\xspace}

@@

\begin{frame}{Context}

\centerline{
    \begin{tabular}{>{\vspace{-1.7ex}\raggedleft}p{0.2\textwidth}@{\qquad}p{0.3\textwidth}@{\qquad}p{0.3\textwidth}}
    & \alert{Previous talks} & \alert{My reponse} \tabularnewline
    \toprule
    \alert{Henryk, Klaus, Greg, Stefan, \ldots} & Many results for $f \in \smallscoop$  &
    May we obtain analogous results for $f \in \smallcone$? \tabularnewline[2ex]
    \alert{Houman} & Let's \alert{learn} the appropriate kernel from the function data & Will only work for nice functions in a \smallcone \tabularnewline[2ex]
    \alert{Paul} & Let's \alert{find} the Lipschitz matrix constant \ldots & Yes, lets learn from the function data \tabularnewline[2ex]
    \alert{Klaus} & This adaptive algorithm has \alert{no theory}  & We want to construct adaptive algorithms \alert{with theory} \tabularnewline[2ex]
    \alert{Henryk} & Tractability & Yes! \tabularnewline[2ex]
    \alert{Greg} & POD weights & Yes!
    \end{tabular}}
\end{frame}

@@

    \item<2>[Step 3] Let $f_{\textup{min}} \in \cf$ be the \alert{minimum} norm interpolant of the data $L_1(f), \ldots, L_n(f)$

    \item[Step 4]  \uncover<2>{If $C\bignorm[\cf]{f_{\textup{min}}} \le \rho$ for some preset \alert{inflation} factor, $C$, }then return $A(f,\varepsilon) = \Sapp(f,n)$\uncover<2>{; otherwise, choose $\rho = 2C\bignorm[\cf]{f_{\textup{min}}}$, and go to Step 2}
\end{description}

@@

\uncover<2>{This succeeds for the \alert{cone} \smallcone defined as those functions in $\cf$ whose norms are not much larger than their minimum norm interpolants.}

@@

\begin{align*}
    \only<1>{\cf &:= \left \{ f = \sum_{\vj \in \cn} \hf(\vj) u_{\vj} : \norm[\cf]{f} := \norm[2]{\left(\frac{\bigabs{\hf(\vj)}}{\lambda_{\vj}} \right)_{\vj \in \cn}} \right \} \qquad \begin{minipage}{3cm}\raggedright\alert{$\vlambda$ affects convergence rate \& tractability}\end{minipage}\\
    \cg &: = \biggl \{ g = \sum_{\vj \in \cn} \hg(\vj) v_{\vj} : \norm[\cg]{g} := \bignorm[2]{\hg}\biggr \}, \qquad v_{\vj} = S(u_{\vj}) \\[-2ex]}
      & \alert{\lambda_{\vj_1} \ge \lambda_{\vj_2} \ge \cdots}, \qquad
      n_0 < n_1 < n_2 < \cdots,  \qquad
 	\sigma_k(f) = \sum_{i=n_{k-1}+1}^{n_k}  \bigabs{\hf(\vj_i)}, \quad k \in \naturals \\
     \cc &: = \left\{f \in \cf :  
 	 \sigma_\ell(f) \le a b^{\ell - k}\sigma_k(f) \quad \forall k \le \ell \right\} \quad a > 1 > b \quad \text{\alert{series coef.\ decay steadily}}\\
 	 \Sapp(f,n_k) &= \sum_{i=1}^{n_k} \hf(\vj_i) v_{\vj_i} \text{ is optimal for fixed }n_k, \qquad \norm[\cg]{S(f) - \Sapp(f,n_k)} \le \frac{ab \sigma_k(f)}{\sqrt{1 - b^2}} \\[-1ex]
 	 \alert{A(f,\varepsilon)} & \alert{= \Sapp(f,n_k)} \text{ for the smallest $k$ satisfying } \alert{\sigma_k(f) \le \frac{\varepsilon \sqrt{1 - b^2}}{ab}}
 	 \only<2>{\\ 
 	 \MoveEqLeft{\alert{\cost(A,\cc,\varepsilon,\rho) = n_{\ell^\dagger}}, \quad 
 	 \ell^\dagger \le \min \left \{\ell \in \naturals : \frac{\rho^2}{\varepsilon^2} \le \frac{(1 - b^2)}{a^2b^2} \left[ \sum_{k=1}^{\ell-1} \frac{b^{2(k-\ell)}}{a^2\lambda_{n_{k-1}+1}^2} + \frac{1}{\lambda_{n_{\ell-1}+1}^2}\right]   \right\}} \\
 	 \MoveEqLeft{\cost(A,\cc,\varepsilon,\rho) \text{ \alert{essentially no worse than} } \comp(\ca(\cc,\LambdaAll),\varepsilon,\rho)}\\
 	 \MoveEqLeft{\text{\alert{No} tractability results}}}
\end{align*}

@@

    \begin{align*}
    \only<1>{\cf &:= \left \{ f = \sum_{\vj \in \natzero^d} \hf(\vj) u_{\vj} : \norm[\cf]{f} :=  \norm[\infty]{\left(\frac{\bigabs{\hf(\vj)}}{\lambda_{\vj}} \right)_{\vj \in \cn}} \right \} \qquad \begin{minipage}{3cm}\raggedright\alert{$\vlambda$ affects convergence rate \& tractability}\end{minipage}\\
    \cg &: = \Biggl \{ g = \sum_{\vj \in \natzero^d} \hg(\vj) v_{\vj} : \norm[\cg]{g} := \bignorm[1]{\hg}\Biggr \}, \qquad v_{\vj} = S(u_{\vj}) = u_{\vj} \\[-2ex]}
    \lambda_{\vj} & = \Gamma_{\norm[0]{\vj}} \prod_{\substack{\ell = 1 \\ j_{\ell} > 0}}^d \gamma_\ell s_{j_\ell} \quad
    \begin{cases}
\gamma_\ell = \text{coordinate importance} \\[-0.5ex]
\Gamma_r = \text{order size} \\[-0.5ex]
s_j =  \text{smoothness degree}
\end{cases}
\text{\alert{POSD weights\footfullcite{WuHam00,KuoEtal12a}}} \\
      & \alert{\lambda_{\vj_1} \ge \lambda_{\vj_2} \ge \cdots}, \qquad
      \Sapp(f,n) = \sum_{i=1}^{n} \hf(\vj_i) u_{\vj_i} ,  \qquad
 	\norm[\cg]{S(f) - \Sapp(f,n_k)} \le \norm[\cf]{f} \sum_{i = n+1} \lambda_{\vj_i} \\
     \cc &: = \left\{f \in \cf :  \text those functions \right \}
 	 \only<2>{\\
 	 \alert{A(f,\varepsilon)} & \alert{= \Sapp(f,n_k) \text{ for the smallest $k$ satisfying } \sigma_k(f) \le \frac{\varepsilon \sqrt{1 - b^2}}{ab}} \\ 
 	 \MoveEqLeft{\alert{\cost(A,\cc,\varepsilon,\rho) = n_{\ell^\dagger}, \quad 
 	 \ell^\dagger \le \min \left \{\ell \in \naturals : \frac{\rho^2}{\varepsilon^2} \le \frac{(1 - b^2)}{a^2b^2} \left[ \sum_{k=1}^{\ell-1} \frac{b^{2(k-\ell)}}{a^2\lambda_{n_{k-1}+1}^2} + \frac{1}{\lambda_{n_{\ell-1}+1}^2}\right]   \right\}}} \\
 	 \MoveEqLeft{\alert{\cost(A,\cc,\varepsilon,\rho) \text{ essentially no worse than } \comp(\ca(\cc,\LambdaAll),\varepsilon,\rho)}}}
\end{align*}

@@

\begin{frame}{Take Home Message}

\begin{itemize}
    \item Assuming that the input functions lie in \alert{cones} \smallcone allow us to construct \alert{adaptive} algorithms

    \item Cone definitions reflect \alert{prior beliefs} or \alert{practical considerations}

    \item Much to be done to 
    \begin{itemize}
        \item demonstrate \alert{solvability}

        \item \alert{construct} algorithms

        \item derive upper bounds on \alert{computational cost}

        \item derive lower bounds on \alert{computational complexity}

        \item determine conditions on \alert{tractability}

        \item \alert{implement} algorithms in available software
    \end{itemize}

@@
